---
title: "Class Notes"
author: "Erica Dale"
date: "2022-08-11"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## For loop practice

```{r}
quarter_splits <- c(1.0, 1.1, 1.2, 1.1, 1.4, 1.5, 1.6, 1.4)

# Return total time for any two consecutive quarter miles (i.e. 1.0 + 1.1)
for(i in seq_along(quarter_splits)){
  half_time <- quarter_splits[i] + quarter_splits[i + 1]
  print(half_time)
}

# Avoid final NA: While loop is not NA, print this
  # Or add an if statement
```

## Intro to Functions

Make a function to double the user input

```{r}
double_this <- function(your_value) {
  print(2 * your_value)
}

double_this(your_value = 5)
double_this(7)

# Automatically vectorized
double_this(c(1,2,10,75))
```

Write function asks for user's NAME and AGE, returns "I'm [NAME] and I'm
[AGE] years old"

```{r}
name_age <- function(your_name, your_age) {
  print(paste("I'm", your_name, "and I'm", your_age, "years old"))
}

name_age("Brit", 25)

```

### Function documentation

```{r}
?mean
?t.test
```

top: name and {package} Title of function should be descriptive
Desciption: Usage: structure and required input arguments, default
settings Arguments: summarize of default and alternatives Value: ideas
of what output names are in stored list Examples: generally not clear,
use google instead

### Function with Conditional

Calculate animal age based on human years

```{r}
convert_animal_age <- function(animal_species, human_years) {
  if (animal_species == "dog") {
    print(human_years * 7)
  } else if (animal_species == "goat") {
    print(human_years * 4.7)
  } else
    print("This animal cannot be calculated")
}

convert_animal_age(animal_species = "dog", human_years = 12)
convert_animal_age("goat", 4)
convert_animal_age("cat", 11)
```


### Function with data wrangling
Make a data frame
```{r}
dog_choice <- tribble(
  ~dog_name, ~food,
  "Khora", "everything",
  "Teddy", "salmon",
  "Waffle", "pancakes",
  "Banjo", "chicken"
)
view(dog_choice)
```

Function given a dogs name, pull information from data frame of food preference
```{r}
# Create subset so only keep row with dog specified
dog_menu <- function(input_dog_name) {
  my_subset <- filter(dog_choice, dog_name == input_dog_name) # Returns row matching
  print(paste("My name is", my_subset$dog_name, "and my favorite food is", my_subset$food))
  }
dog_menu("Waffle")

# OR
dog_menu2 <- function(input_dog_name) {
  dog_choice |> 
  filter(dog_name == input_dog_name)
  print(paste("My name is", my_subset$dog_name, "and my favorite food is", my_subset$food)) 
  }
dog_menu2("Banjo") # DOES NOT WORK, Where did it break
```


### Functions and For Loops
Function that iterates over data in given df, calculates mean

```{r}
means_over_columns <- function(df) {
  for(i in seq_along(df)) {
    col_mean <- mean(df[[i]], na.rm = TRUE)  # intermediate to hold the calculated data
    col_name <- colnames(df[i])  # specific function to pull out column names
    print(paste("The mean value of", col_name, "is", col_mean))
  }
}

means_over_columns(df = mtcars)
means_over_columns(df = faithful)
# Also can use colmeans, apply, across, plus more as built in functions for this
```


#### Nested For Loop

```{r}
plants <- c("orchid", "philodendron", "snake plant")
rooms <- c("bedroom", "living room", "office", "kitchen")

plant_rooms <- matrix(nrow = length(plants), ncol = length(rooms))
for (i in seq_along(rooms)) {
  for(j in seq_along(plants)) {
  organized_rooms <- print(paste(rooms[i], plants[j]))
  plant_rooms[j, i] <- organized_rooms
  }
}
plant_rooms
```


## Logistic Growth Equation
Models population growth as it nears a carrying capacity
r= growth rate
K= carrying capacity

#### First with for loop
1. What do I want? automate estimating population size over a range of times at multiple growth rates (for a constant initial population and carrying capacity)
Which variables as columns vs rows (time and growth rate)? EITHER is fine, just keep track

2. Outline: Want to write out how I think this could code
Pseudocode:
Function logistic_growth, iterate over multiple times for multiple growth rates
Create empty shell
Nested for loop over TIME then GROWTH

out_matrix = matrix() # to store output
for(i in sequence of growth rates) {
for(j in sequence of times) {
calc_logistic(for i over all j)
store calc_population in out_matrix
}
}

row = time, column = growth rate

Bare bones function and test:
```{r}
logistic_growth <- function(N0, K, r, time) {
  Nt <- K / (1 + ((K - N0) / N0) * exp(-r * time))
  print(Nt)
}
```

Test it
```{r}
time_vec <- seq(from = 0, to = 35, by = 0.1)

pop_35_vec <- vector(mode = "numeric", length = length(time_vec))

for(i in 1:length(time_vec)) {
  population <- logistic_growth(N0 = 100, K = 6000, r = .27, time = time_vec[i])
  pop_35_vec <- population
}

```

Want to apply it wider
```{r}
r_seq <- seq(from = .2, to = .4, by = .1)
time <- (60, 80, 180)

out_matrix <- matrix(nrow = length(time_vec), ncol = length(r_seq))

for (i in 1:length(r_seq)) {
  for (j in 1:length(time_vec)) {
    pop <- logistic_growth(N0 = 100, K = 6000, r = r_seq[i], time = time_vec[j])
    out_matrix[j, i] <- pop
  }
}

view(out_matrix)
# messed up something?
```


## Lotistic growth Function
Walkthrough in class, properly done (unlike aboves attempt)

```{r}
# Write out the equation
K / (1 + ((K - N0) / N0) * exp(-r * time))

# Shortcut ctrl-alt-x
logistic_growth <- function(K, N0, r, time) {
  K / (1 + ((K - N0) / N0) * exp(-r * time))
}

# Test
logistic_growth(K = 5000, N0 = 100, r = 0.07, time = 42)
```

Add a time vector to iterate over time
```{r}
time_vec <- seq(from = 0, to = 35, by = .1)
time_vec

pop_35 <- logistic_growth(N0 = 40, K = 4000, r = .4, time = time_vec)
glimpse(pop_35)
```

For plotting we want data together (df) instead of separate vectors
```{r}
pop_time35_df <- data.frame(time_vec, pop_35)
# View the data!

ggplot(data = pop_time35_df, aes(x = time_vec, y = pop_35)) +
  geom_line()

```


Let's do this ONE OTHER WAY
Instead of a vector that it auto-goes over, put it in a for loop
```{r}
pop_35_vector <- vector(mode = "numeric", length = length(time_vec))

for (i in seq_along(time_vec)) {
  population <- logistic_growth(N0 = 100, K = 6000, r = 0.27, time = time_vec[i])
  pop_35_vector[i] <- population
}
```

Now extend this to iterate over multiple growth rates
```{r}
r_seq <- seq(from = 0.2, to = 0.4, by = 0.01)
# Check code chunk in consol
```

```{r}
out_m <- matrix(nrow = length(time_vec), ncol = length(r_seq))

# What is highest level iteration? Across the growth rates
for (i in seq_along(r_seq)) {
  for (j in seq_along(time_vec)) {
    pop <- logistic_growth(N0 = 100, K = 1000, r = r_seq[i], time = time_vec[j])
    out_m[j, i] <- pop
  }
}
```

Now some wrangling of the result data
1. bind together out matrix with new label column
2. add description to columns

```{r}
logistic_df <- data.frame(out_m, time = time_vec) # add time column

#Update column names
colnames(logistic_df) <- c(paste0("gr_", r_seq), "time")  # time column for last column
```

BUT NOT IN TIDY DATA STRUCTURE!
three variables: population (split up over mulitple columns), growth rate (does not have it's own column), and time

```{r}
logistic_df_long <- logistic_df |> 
  pivot_longer(cols = -time, values_to = "population", # time is already set
                names_to = "growth_rate") # keep column names that have growth rates
```

Now graph it!
```{r}
ggplot(data = logistic_df_long, aes(x = time, y = population)) +
  geom_line()
# Multiple population values at each time, so high frequency jagged line
```

Variable to group by:
```{r}
ggplot(data = logistic_df_long, aes(x = time, y = population)) +
  geom_line(aes(color = growth_rate))
```



# Warnings and Error Message
use `warning` to add a warning
use `stop` to add an error

Collective wind turbine power

$$
P=0.3\rho A v^3
$$

rho: air density (kg/ m^3) - 1.225 is at sea level
A: area covered by turbine (square meters)
v: wind velocity (m/s)
result in  watts?

User inputs radius of turbine
```{r}
# write it out
0.3*rho*pi*(radius^2)*(windspeed^3)

# shortcut to create function: highlight and cmd-alt-x
calc_windpower <- function(rho, radius, windspeed) {
  
  #add warnings
  if (windspeed > 130) {
    warning("WHOA, that's really fast. Are you sure?")
  }
  if (rho > 1.225) {
    warning("That air density is suspicious.")
  }
  if (radius < 0) {
    stop("Rotor radius must be a positive value in meters.")
  }
  
  0.3*rho*pi*(radius^2)*(windspeed^3)
}

calc_windpower(1.5, 50, 35)
```

