---
title: "Class Notes"
author: "Erica Dale"
date: "2022-08-11"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## For loop practice

```{r}
quarter_splits <- c(1.0, 1.1, 1.2, 1.1, 1.4, 1.5, 1.6, 1.4)

# Return total time for any two consecutive quarter miles (i.e. 1.0 + 1.1)
for(i in seq_along(quarter_splits)){
  half_time <- quarter_splits[i] + quarter_splits[i + 1]
  print(half_time)
}

# Avoid final NA: While loop is not NA, print this
  # Or add an if statement
```

## Intro to Functions

Make a function to double the user input

```{r}
double_this <- function(your_value) {
  print(2 * your_value)
}

double_this(your_value = 5)
double_this(7)

# Automatically vectorized
double_this(c(1,2,10,75))
```

Write function asks for user's NAME and AGE, returns "I'm [NAME] and I'm
[AGE] years old"

```{r}
name_age <- function(your_name, your_age) {
  print(paste("I'm", your_name, "and I'm", your_age, "years old"))
}

name_age("Brit", 25)

```

## Function documentation

```{r}
?mean
?t.test
```

top: name and {package} Title of function should be descriptive
Desciption: Usage: structure and required input arguments, default
settings Arguments: summarize of default and alternatives Value: ideas
of what output names are in stored list Examples: generally not clear,
use google instead

## Function with Conditional

Calculate animal age based on human years

```{r}
convert_animal_age <- function(animal_species, human_years) {
  if (animal_species == "dog") {
    print(human_years * 7)
  } else if (animal_species == "goat") {
    print(human_years * 4.7)
  } else
    print("This animal cannot be calculated")
}

convert_animal_age(animal_species = "dog", human_years = 12)
convert_animal_age("goat", 4)
convert_animal_age("cat", 11)
```


## Function with data wrangling
Make a data frame
```{r}
dog_choice <- tribble(
  ~dog_name, ~food,
  "Khora", "everything",
  "Teddy", "salmon",
  "Waffle", "pancakes",
  "Banjo", "chicken"
)
view(dog_choice)
```

Function given a dogs name, pull information from data frame of food preference
```{r}
# Create subset so only keep row with dog specified
dog_menu <- function(input_dog_name) {
  my_subset <- filter(dog_choice, dog_name == input_dog_name) # Returns row matching
  print(paste("My name is", my_subset$dog_name, "and my favorite food is", my_subset$food))
  }
dog_menu("Waffle")

# OR
dog_menu2 <- function(input_dog_name) {
  dog_choice |> 
  filter(dog_name == input_dog_name)
  print(paste("My name is", my_subset$dog_name, "and my favorite food is", my_subset$food)) 
  }
dog_menu2("Banjo") # DOES NOT WORK, Where did it break
```


## Functions and For Loops
Function that iterates over data in given df, calculates mean

```{r}
means_over_columns <- function(df) {
  for(i in seq_along(df)) {
    col_mean <- mean(df[[i]], na.rm = TRUE)  # intermediate to hold the calculated data
    col_name <- colnames(df[i])  # specific function to pull out column names
    print(paste("The mean value of", col_name, "is", col_mean))
  }
}

means_over_columns(df = mtcars)
means_over_columns(df = faithful)
# Also can use colmeans, apply, across, plus more as built in functions for this
```

